\begin{definition}[Per-coordinate DP-noise variance]\label{def:phi}
Fix a trainable coordinate of the DP-LoRA model. Following the DP-Adam update populated by Opacus (see \texttt{src/dp\_optim/dp\_adaptive.py}, the \texttt{phi} property), the privatized minibatch gradient seen by the optimizer is
\[
g_t \;=\; \bar g_t + \xi_t,\qquad \xi_t \sim \mathcal N(0,\Phi),\qquad
\Phi \;:=\; \Big(\frac{\sigma_{\mathrm{DP}}\,C}{B_{\mathrm{eff}}}\Big)^2,
\]
where $\bar g_t=\tfrac1{B_{\mathrm{eff}}}\sum_{i\in\mathcal B_t}\operatorname{clip}_C(\nabla\ell_i)$ is the clipped, batch-averaged (conditionally deterministic) gradient mean, $\sigma_{\mathrm{DP}}$ is the noise multiplier, $C$ the clipping norm, and $B_{\mathrm{eff}}=\texttt{expected\_batch\_size}\cdot\texttt{accumulated\_iterations}\,(\cdot\,\texttt{world\_size})$ the effective batch size. The DP noise $\xi_t$ is drawn independently of $\bar g_t$ and is zero-mean with the same variance $\Phi$ on every coordinate.
\end{definition}

\begin{theorem}[The $\rho$-diagnostic for DP-AdamBC]\label{thm:rho}
Let $\hat v$ denote Adam's bias-corrected second-moment estimate at the coordinate, and let
\[
v_{\mathrm{true}} \;:=\; \mathbb E[\hat v]-\Phi
\]
be its noise-free counterpart, with $v_{\mathrm{true}}\ge 0$ and $\Phi>0$. Define the \emph{noise share}
\[
\rho \;:=\; \frac{\Phi}{\mathbb E[\hat v]} \;=\; \frac{\Phi}{v_{\mathrm{true}}+\Phi}\;\in\;(0,1].
\]
Then:
\begin{enumerate}[leftmargin=2.4em,labelsep=0.4em,itemsep=4pt]
\item[\textnormal{(i)}] \textbf{(Inflation / debiasing)}\enspace  $\mathbb E[\hat v]=v_{\mathrm{true}}+\Phi$, so the DP-AdamBC corrected denominator recovers the signal second moment in expectation: $\mathbb E[\hat v-\Phi]=v_{\mathrm{true}}$.
\item[\textnormal{(ii)}] \textbf{(Step-correction factor)}\enspace  In expectation (in the ratio-of-expected-denominators sense made precise below) the multiplicative change DP-AdamBC applies to the per-coordinate Adam step is
\[
\kappa_{\mathrm{BC}}(\rho)\;:=\;\frac{\sqrt{\mathbb E[\hat v]}}{\sqrt{v_{\mathrm{true}}}}\;=\;\frac1{\sqrt{1-\rho}},
\]
with relative change $g(\rho):=\kappa_{\mathrm{BC}}(\rho)-1=(1-\rho)^{-1/2}-1$.
\item[\textnormal{(iii)}] \textbf{(Materiality)}\enspace  Writing $r:=\Phi/v_{\mathrm{true}}=\rho/(1-\rho)\ge0$, the correction is an \emph{order-one} relative change iff $r=\Theta(1)$, i.e.\ iff $v_{\mathrm{true}}=\Theta(\Phi)$, equivalently iff $\rho$ is bounded away from both $0$ and $1$. Quantitatively $g(\rho)$ is sandwiched, \emph{for all $r\ge 0$}, as
\[
\tfrac12\,r \;-\;\tfrac12 r^2\;\le\;g(\rho)\;\le\;\tfrac12\, r ,\qquad r=\frac{\rho}{1-\rho},
\]
so $g(\rho)\to 0$ as $\rho\to 0$ (BC inert) and $g(\rho)\to\infty$ as $\rho\to 1$.
\item[\textnormal{(iv)}] \textbf{(Maximal sensitivity / sweet spot)}\enspace  On the natural scale $u:=\log r=\log(\Phi/v_{\mathrm{true}})$ the noise share $\rho(u)=\dfrac{e^{u}}{1+e^{u}}$ is the logistic sigmoid; its sensitivity $\dfrac{d\rho}{du}=\rho(1-\rho)$ attains its unique maximum $\tfrac14$ at
\[
\rho=\tfrac12 \iff \Phi=v_{\mathrm{true}} \iff r=1 .
\]
Thus $\rho=\tfrac12$ is the diagnostic's most sensitive operating point: the crossover where the DP-noise variance exactly equals the true second moment.
\item[\textnormal{(v)}] \textbf{(Saturation)}\enspace  $\rho\to 1$ monotonically as $v_{\mathrm{true}}\to 0^{+}$; equivalently $\rho$ is strictly decreasing in $v_{\mathrm{true}}$ for fixed $\Phi$, with $\rho=1$ attained only in the limit $v_{\mathrm{true}}=0$.
\end{enumerate}
\end{theorem}

\begin{proof}
Throughout, fix one trainable coordinate; all quantities are scalar. We use Definition~\ref{def:phi}: $g_t=\bar g_t+\xi_t$ with $\xi_t\sim\mathcal N(0,\Phi)$ independent of the clipped mean $\bar g_t$, and $\Phi=(\sigma_{\mathrm{DP}}C/B_{\mathrm{eff}})^2>0$. Adam maintains the exponential moving average $v_t=\beta_2 v_{t-1}+(1-\beta_2)g_t^2$ ($v_0=0$) with bias correction $\hat v_t=v_t/(1-\beta_2^{\,t})$, exactly as implemented in \texttt{DPAdaptive.\_adam\_update}. We suppress the time index $t$ where it is inessential and write $\mathbb E[\hat v]$ for the conditional expectation over the DP noise $\{\xi_s\}_{s\le t}$ given the (deterministic) clipped means $\{\bar g_s\}_{s\le t}$; the unconditional statement then follows by the tower property, since $\Phi$ is a fixed constant and the quantities below are affine in $\{\bar g_s^2\}$.

\medskip
\noindent\textbf{(i) Inflation and exact debiasing in expectation.}
We first establish the inflation identity $\mathbb E[\hat v]=v_{\mathrm{true}}+\Phi$ with
\[
v_{\mathrm{true}}\;:=\;\frac{(1-\beta_2)\sum_{k=0}^{t-1}\beta_2^{k}\,\bar g_{t-k}^2}{1-\beta_2^{\,t}} ,
\]
the bias-corrected EWMA of the \emph{clipped, noise-free} squared gradients (note $v_{\mathrm{true}}\ge0$ since every summand is nonnegative). Conditioning on $\bar g_s$ and using $\mathbb E[\xi_s]=0$, $\mathbb E[\xi_s^2]=\Phi$, and the independence of $\xi_s$ from $\bar g_s$,
\[
\mathbb E\!\left[g_s^2\mid \bar g_s\right]
=\mathbb E\!\left[(\bar g_s+\xi_s)^2\mid \bar g_s\right]
=\bar g_s^2 + 2\bar g_s\,\mathbb E[\xi_s] + \mathbb E[\xi_s^2]
=\bar g_s^2+\Phi,
\]
because the cross term $2\bar g_s\,\mathbb E[\xi_s]=0$ vanishes. Unrolling the linear EWMA recursion from $v_0=0$ gives $v_t=(1-\beta_2)\sum_{k=0}^{t-1}\beta_2^{k}g_{t-k}^2$, hence by linearity of expectation
\[
\mathbb E[v_t]
=(1-\beta_2)\sum_{k=0}^{t-1}\beta_2^{k}\big(\bar g_{t-k}^2+\Phi\big)
=(1-\beta_2)\sum_{k=0}^{t-1}\beta_2^{k}\,\bar g_{t-k}^2 \;+\;\Phi\,(1-\beta_2)\sum_{k=0}^{t-1}\beta_2^{k}.
\]
The geometric sum evaluates to $(1-\beta_2)\sum_{k=0}^{t-1}\beta_2^{k}=(1-\beta_2)\cdot\frac{1-\beta_2^{\,t}}{1-\beta_2}=1-\beta_2^{\,t}$. Dividing by the bias-correction factor $1-\beta_2^{\,t}$ gives
\[
\mathbb E[\hat v_t]=\frac{\mathbb E[v_t]}{1-\beta_2^{\,t}}
= v_{\mathrm{true}} + \Phi .
\]
The crucial point is that after bias correction the additive contamination is the \emph{constant} $\Phi$ (the factor $1-\beta_2^{\,t}$ has cancelled), which is exactly what makes a constant subtraction the correct debiasing. DP-AdamBC (\citealp{tang2024dpadambc}; \texttt{bias\_correction=True} in \texttt{\_adam\_update}, where $\hat v\leftarrow\max(\hat v-\Phi,\xi)$) therefore satisfies, whenever the floor clamp is inactive,
\[
\mathbb E\big[\hat v - \Phi\big]=\mathbb E[\hat v]-\Phi=(v_{\mathrm{true}}+\Phi)-\Phi=v_{\mathrm{true}},
\]
i.e.\ the corrected denominator recovers the signal second moment $v_{\mathrm{true}}$ in expectation. This proves (i). In particular $\rho=\Phi/\mathbb E[\hat v]=\Phi/(v_{\mathrm{true}}+\Phi)$ is well defined and, since $v_{\mathrm{true}}\ge 0$ and $\Phi>0$, lies in $(0,1]$, with $\rho=1$ iff $v_{\mathrm{true}}=0$.

\medskip
\noindent\textbf{(ii) The step-correction factor $\kappa_{\mathrm{BC}}(\rho)=(1-\rho)^{-1/2}$.}
Adam's per-coordinate update has magnitude proportional to $\hat m/(\sqrt{\hat v}+\epsilon)$ (\texttt{p.addcdiv\_(mhat, vhat.sqrt().add\_(eps), -lr)}). With $\hat m$ held fixed (BC changes only the denominator) and the regularizer $\epsilon$ negligible against $\sqrt{\hat v}$, we compare the \emph{deterministic} surrogate steps obtained by replacing the random denominator $\sqrt{\hat v}$ by the square root of its expectation, $\sqrt{\mathbb E[\hat v]}$ (the inflated denominator) versus $\sqrt{\mathbb E[\hat v]-\Phi}=\sqrt{v_{\mathrm{true}}}$ (the debiased denominator). Their ratio is
\[
\kappa_{\mathrm{BC}}
:=\frac{\text{step}_{\mathrm{BC}}}{\text{step}_{\mathrm{no\,BC}}}
=\frac{\sqrt{\mathbb E[\hat v]}}{\sqrt{\mathbb E[\hat v]-\Phi}}
=\sqrt{\frac{v_{\mathrm{true}}+\Phi}{v_{\mathrm{true}}}}
=\sqrt{\frac{1}{1-\rho}},
\]
where the last equality uses $v_{\mathrm{true}}=(1-\rho)\,\mathbb E[\hat v]$, which follows from $\rho=\Phi/\mathbb E[\hat v]$ and $v_{\mathrm{true}}=\mathbb E[\hat v]-\Phi$. Hence $\kappa_{\mathrm{BC}}(\rho)=(1-\rho)^{-1/2}$ and $g(\rho)=\kappa_{\mathrm{BC}}(\rho)-1=(1-\rho)^{-1/2}-1$. This is the amount by which DP-AdamBC enlarges the noise-shrunk Adam step at the level of expected denominators. (At the empirically reachable point $\rho=0.955$ this gives $\kappa_{\mathrm{BC}}=1/\sqrt{1-0.955}=4.71$, consistent with the measured $\approx 3.8$--$4.7\times$ effective-step increase; as $\rho\to 1$ it diverges, reflecting the unstable over-subtraction regime.)

\emph{Scope of the claim.} $\kappa_{\mathrm{BC}}$ is defined as the ratio of the square roots of the \emph{expected} denominators, i.e.\ a plug-in (first-moment / delta-method) approximation to the random per-step ratio $\sqrt{\hat v}/\sqrt{\hat v-\Phi}$. This is precisely the quantity the project's telemetry measures (medians of $\sqrt{\hat v}$). For the strictly stochastic per-step ratio one would invoke Jensen: since $x\mapsto 1/\sqrt{x}$ is convex on $(0,\infty)$, $\mathbb E[1/\sqrt{\hat v-\Phi}]\ge 1/\sqrt{\mathbb E[\hat v-\Phi]}=1/\sqrt{v_{\mathrm{true}}}$, so the true expected enlargement is \emph{at least} $\kappa_{\mathrm{BC}}$; the plug-in value is a lower bound and the Jensen gap only sharpens the $\rho\to1$ instability conclusion. This proves (ii) in the stated sense.

\medskip
\noindent\textbf{(iii) Materiality: order-one change iff $v_{\mathrm{true}}=\Theta(\Phi)$.}
Substitute $\rho=r/(1+r)$, so $1-\rho=1/(1+r)$ and $\kappa_{\mathrm{BC}}=\sqrt{1+r}$, hence
\[
g(\rho)=\sqrt{1+r}-1,\qquad r=\frac{\Phi}{v_{\mathrm{true}}}=\frac{\rho}{1-\rho}\ \ge 0 .
\]
We bound $h(r):=\sqrt{1+r}-1$ on $[0,\infty)$.

\emph{Upper bound.} The map $r\mapsto\sqrt{1+r}$ is concave on $[0,\infty)$, so it lies below its tangent line at $r=0$: $\sqrt{1+r}\le 1+\tfrac12 r$. Subtracting $1$,
\[
g(\rho)=\sqrt{1+r}-1\;\le\;\tfrac12\,r\qquad(\forall\,r\ge0).
\]

\emph{Lower bound.} We claim $g(\rho)=\sqrt{1+r}-1\ge \tfrac12 r-\tfrac12 r^2$ for \emph{all} $r\ge 0$, and we split into two cases.

\noindent\textit{Case $r>1$.} Then $\tfrac12 r-\tfrac12 r^2=\tfrac12 r(1-r)<0\le g(\rho)$, so the bound holds trivially.

\noindent\textit{Case $0\le r\le 1$.} Set $a:=1+\tfrac12 r-\tfrac12 r^2$. On $[0,1]$ we have $a=1+\tfrac12 r(1-r)\ge 1>0$. It therefore suffices to show $a^2\le 1+r$ (then $a\le\sqrt{1+r}$, i.e.\ $g(\rho)=\sqrt{1+r}-1\ge a-1=\tfrac12 r-\tfrac12 r^2$). Expanding,
\[
a^2=\Big(1+\tfrac12 r-\tfrac12 r^2\Big)^2=1+r-\tfrac34 r^2-\tfrac12 r^3+\tfrac14 r^4,
\]
so
\[
(1+r)-a^2=\tfrac34 r^2+\tfrac12 r^3-\tfrac14 r^4=\frac{r^2}{4}\,(3+2r-r^2)=\frac{r^2}{4}\,(3-r)(1+r)\;\ge\;0
\]
for all $r\in[0,3]$, in particular on $[0,1]$. (This identity makes precise where the squared inequality is valid: exactly for $r\le 3$; we only need $r\le1$.) Hence $a^2\le 1+r$ on $[0,1]$, giving the lower bound there.

Combining the two cases with the upper bound,
\[
\tfrac12 r-\tfrac12 r^2\;\le\;g(\rho)\;\le\;\tfrac12 r\qquad(\forall\,r\ge0).
\]
In particular, for bounded $r$ the two bounds pinch $g(\rho)=\tfrac12 r+O(r^2)$, so $g(\rho)=\Theta(r)$, and:
\begin{itemize}
\item $g(\rho)=\Theta(1)$ (a genuine order-one change to the step) \emph{iff} $r=\Theta(1)$, i.e.\ $v_{\mathrm{true}}=\Theta(\Phi)$, i.e.\ $\rho$ bounded away from $0$ and $1$;
\item $g(\rho)\le \tfrac12 r\to 0$ as $\rho\to 0$ ($\Phi\ll v_{\mathrm{true}}$): DP-AdamBC and DP-Adam coincide to first order, so bias correction is a no-op;
\item $g(\rho)=(1-\rho)^{-1/2}-1\to\infty$ as $\rho\to 1$ ($v_{\mathrm{true}}\ll\Phi$): the correction is unboundedly large, the regime where $\hat v-\Phi\to 0$ and the floor $\xi$ takes over (the ``collapse to momentum-SGD'' regime documented in the experiments).
\end{itemize}
This proves (iii): bias correction is material exactly when $v_{\mathrm{true}}=\Theta(\Phi)$. (The earlier project bound $|\kappa_{\mathrm{BC}}-1|\le r/2$ in \texttt{theory.tex} is exactly the upper half established here.)

\medskip
\noindent\textbf{(iv) Maximal sensitivity at $\rho=\tfrac12$.}
The materiality threshold $r=\Theta(1)$ singles out a multiplicative scale; the canonical coordinate is $u:=\log r=\log(\Phi/v_{\mathrm{true}})$, the log noise-to-signal ratio (well defined for $v_{\mathrm{true}}>0$, i.e.\ $\rho\in(0,1)$). From $\rho=r/(1+r)$ with $r=e^{u}$,
\[
\rho(u)=\frac{e^{u}}{1+e^{u}}=\frac{1}{1+e^{-u}},
\]
the logistic sigmoid, which is smooth and strictly increasing from $\rho(-\infty)=0$ to $\rho(+\infty)=1$. Its sensitivity is
\[
\frac{d\rho}{du}=\frac{e^{u}(1+e^{u})-e^{u}\cdot e^{u}}{(1+e^{u})^2}=\frac{e^{u}}{(1+e^{u})^2}=\rho(u)\,\big(1-\rho(u)\big).
\]
Maximizing $s(\rho):=\rho(1-\rho)$ over $\rho\in(0,1)$: $s'(\rho)=1-2\rho=0\iff\rho=\tfrac12$, and $s''(\rho)=-2<0$ confirms a strict interior maximum, with value $s(\tfrac12)=\tfrac12\cdot\tfrac12=\tfrac14$. Equivalently, $\tfrac{d^2\rho}{du^2}=\tfrac{d}{du}\big[\rho(1-\rho)\big]=(1-2\rho)\tfrac{d\rho}{du}=\rho(1-\rho)(1-2\rho)=0\iff\rho=\tfrac12$, the inflection point of the sigmoid. At this point
\[
\rho=\tfrac12\iff r=e^{u}=1\iff \Phi=v_{\mathrm{true}} .
\]
Hence $\rho=\tfrac12$ is the unique operating point of maximal diagnostic sensitivity: a small change in the log noise/signal ratio produces the largest change in $\rho$, and it is exactly the crossover where the DP-noise variance equals the true second moment $\Phi=v_{\mathrm{true}}$. This is the ``sweet spot'': for $\rho<\tfrac12$ the noise is sub-dominant and BC is a small correction ($g\le\tfrac12 r<\tfrac12$); for $\rho>\tfrac12$ the noise dominates and the correction $g$ grows past $1$ and then diverges as $\rho\to1$, so the well-behaved order-one regime is centered at $\rho=\tfrac12$. This proves (iv).

\medskip
\noindent\textbf{(v) Saturation $\rho\to 1$ as $v_{\mathrm{true}}\to 0$.}
With $\Phi>0$ fixed, view $\rho=\Phi/(v_{\mathrm{true}}+\Phi)$ as a function of $v_{\mathrm{true}}\ge 0$. Then
\[
\frac{\partial\rho}{\partial v_{\mathrm{true}}}=-\frac{\Phi}{(v_{\mathrm{true}}+\Phi)^2}<0,
\]
so $\rho$ is strictly decreasing in $v_{\mathrm{true}}$. As $v_{\mathrm{true}}\to 0^{+}$ we get $\rho\to \Phi/\Phi=1$; as $v_{\mathrm{true}}\to\infty$, $\rho\to 0$. Thus $\rho$ saturates monotonically at its ceiling $1$ precisely when the signal second moment falls to zero, $\rho=1$ being attained only in the limit $v_{\mathrm{true}}=0$. This is the regime measured throughout the experiments: $\rho=1.00$--$1.07$ across RoBERTa-large/MNLI ($\varepsilon\in\{1,3,8\}$, $B\in[16,2048]$) and Qwen2.5-1.5B/E2E ($B\le512$), i.e.\ $v_{\mathrm{true}}$ lies below the DP-noise floor, placing the optimizer at $\rho\approx1$ where, by (iii), $g(\rho)$ is large but the corrected denominator $\hat v-\Phi\to0$ and the floor $\xi$ governs the step. The logged ratio saturating at $1$ (rather than at the sweet spot $\tfrac12$) is therefore the diagnostic signature that bias correction cannot act as intended in this recipe.

\smallskip
\noindent\emph{(Measured ratios slightly exceeding $1$.} The reported $\rho\in[1.00,1.07]$ takes values just above the theoretical ceiling $1$; this is consistent with the theory because the logged quantity is $\Phi/\operatorname{median}(\hat v)$ over a finite sample of coordinates rather than $\Phi/\mathbb E[\hat v]$, and a downward sampling/median fluctuation of $\hat v$ below its mean $v_{\mathrm{true}}+\Phi\approx\Phi$ pushes the ratio marginally above $1$. The population quantity $\rho=\Phi/\mathbb E[\hat v]$ is bounded by $1$ as proved.) This proves (v) and completes the proof. \qed

\medskip
\noindent\emph{Remark (consistency of the floored estimator).} The above treats the regime where DP-AdamBC's floor clamp $\hat v\leftarrow\max(\hat v-\Phi,\xi)$ is inactive, which is exactly where the debiasing identity $\mathbb E[\hat v-\Phi]=v_{\mathrm{true}}$ holds. When $\rho\to1$ the residual $\hat v-\Phi$ falls below $\xi$ on a fraction of coordinates approaching $1$ (the measured clamp fraction $\to1$), and the clamp introduces a Jensen gap $\mathbb E[\max(\hat v-\Phi,\xi)]\ge\max(v_{\mathrm{true}},\xi)$; there the estimator is biased upward toward the constant $\xi$ and the update degenerates to $\hat m/\sqrt{\xi}$, i.e.\ momentum-SGD with fixed step scale $1/\sqrt{\xi}$. This is consistent with (iii): the unbounded $g(\rho)$ as $\rho\to1$ is not realized as a useful step but is absorbed by the floor, which is why the maximal-sensitivity sweet spot $\rho=\tfrac12$ --- not the saturation point $\rho=1$ --- is the regime in which bias correction can be both material \emph{and} well-conditioned.

\medskip
\noindent\emph{Remark (privacy-side cross-check of the companion $\kappa$).} The diagnostic above governs the \emph{second}-moment lever (DP-AdamBC). The project's positive-method attempt (DP-CorrMom) instead acts on the \emph{first}-moment / prefix-sum path by injecting anti-correlated noise $w_t=z_t-\lambda z_{t-1}$, whose privacy sensitivity is the distinct scalar $\kappa(\lambda,T)=\operatorname{corr\_sensitivity}(\lambda,T)$ in \texttt{dp\_adaptive.py}. We independently re-derived it to confirm the implemented value. Writing the injection as $n=Lz$ with $L$ lower-bidiagonal ($1$ on the diagonal, $-\lambda$ on the sub-diagonal), the matrix mechanism for the prefix-sum workload $A$ uses strategy $\mathbf C=L^{-1}$, the lower-triangular Toeplitz matrix with entries $(\mathbf C)_{ij}=\lambda^{\,i-j}$ for $i\ge j$ (and $0$ otherwise). The single-participation $\ell_2$ sensitivity is the maximum column norm of $\mathbf C$. Column $j$ has entries $\lambda^{0},\lambda^{1},\dots,\lambda^{T-1-j}$ down the column, so
\[
\|\mathbf C e_j\|_2^2=\sum_{k=0}^{T-1-j}\lambda^{2k},
\]
which is maximized at $j=0$ (the first column, the longest run), giving
\[
\kappa(\lambda,T)=\max_j\|\mathbf C e_j\|_2=\Big(\sum_{k=0}^{T-1}\lambda^{2k}\Big)^{1/2}=\sqrt{\frac{1-\lambda^{2T}}{1-\lambda^{2}}}\;\xrightarrow[T\to\infty]{}\;\frac{1}{\sqrt{1-\lambda^{2}}}.
\]
A Mahalanobis cross-check confirms this: the privatized output is $\hat g\sim\mathcal N(g,\nu^2 LL^{\!\top})$, and a unit change in the data at step $t$ gives Mahalanobis separation $\propto (LL^{\!\top})^{-1}_{tt}=\|(L^{-1})e_t\|_2^2$, again the squared column norm of $L^{-1}$. We verified numerically (e.g.\ $\lambda=0.9$, $T=8$) that the exact maximum column norm of $L^{-1}$ equals $\sqrt{(1-\lambda^{2T})/(1-\lambda^{2})}=2.0707$ and that the large-$T$ limit is $1/\sqrt{1-\lambda^2}=2.2942$. The naive alternative $\sqrt{1+\lambda^2}=1.345$ is the column norm of $L$ \emph{itself} (not $L^{-1}$); using it would \emph{under}-noise the mechanism and \emph{break} the privacy guarantee. Thus the implemented $\kappa=\sqrt{(1-\lambda^{2T})/(1-\lambda^2)}$ is correct, and the alternative $\sqrt{1+\lambda^2}$ is refuted; $\lambda=0$ recovers $\kappa=1$ (i.i.d.\ DP-SGD), as required.
\end{proof}
