\begin{theorem}[DP-CorrMom privacy via the bidiagonal matrix mechanism]
\label{thm:corrmom-privacy}
Fix a number of steps $T\in\mathbb{N}$, a per-sample clipping norm $C>0$, and a correlation
coefficient $\lambda\in[0,1)$. Consider the DP-CorrMom release in which, at each step
$t\in\{1,\dots,T\}$, after per-sample $\ell_2$ clipping at norm $C$ one outputs the noisy
summed gradient
\begin{equation}
\hat g_t \;=\; s_t + n_t,\qquad
s_t := \sum_{i\in\mathcal B_t}\operatorname{clip}_C\!\big(\nabla\ell_i(x_t)\big)\in\mathbb R^{d},
\label{eq:release}
\end{equation}
with the \emph{anti-correlated} Gaussian injection
\begin{equation}
n_t \;=\; z_t-\lambda\,z_{t-1},\qquad
z_0,z_1,\dots,z_T \stackrel{\textnormal{iid}}{\sim}\mathcal N\!\big(0,\nu^2 I_d\big),\quad z_0:=0 .
\label{eq:inject}
\end{equation}
Assume \emph{single participation}: the index sets $\mathcal B_1,\dots,\mathcal B_T$ are disjoint
(every example contributes to at most one step; e.g.\ a single pass with a fixed disjoint partition).
Let $A\in\mathbb R^{T\times T}$ be the lower-triangular all-ones prefix-sum ``workload'' matrix
($A_{t,k}=\mathbf 1[k\le t]$). Then the following hold.
\begin{enumerate}[label=\textnormal{(\roman*)}]
\item \textnormal{(Single Gaussian mechanism.)} Per coordinate, the entire $T$-step release is a
\emph{single fixed-covariance} Gaussian mechanism. Writing the injection as $n=Lz$ with
$L\in\mathbb R^{T\times T}$ lower bidiagonal ($L_{tt}=1$, $L_{t,t-1}=-\lambda$), the prefix-sum
workload $A$ factorises as $A=B\,C_{\mathrm{strat}}$ with \emph{reconstruction} $B:=AL$ and
\emph{strategy} $C_{\mathrm{strat}}:=L^{-1}$ (indeed $(AL)\,L^{-1}=A$), and the released prefix-sum
stream is $A\hat g=As+\nu\,(AL)\,\zeta$ with $\zeta\sim\mathcal N(0,I)$ per coordinate.
Its privacy is governed by a \emph{single} effective noise multiplier
\begin{equation}
\sigma_{\mathrm{eff}}\;=\;\frac{\nu}{C\,\kappa},\qquad
\kappa\;:=\;\max_{1\le t\le T}\big\|(L^{-1})_{:,t}\big\|_2
\;=\;\sqrt{\frac{1-\lambda^{2T}}{1-\lambda^{2}}}\,,
\label{eq:kappa}
\end{equation}
in the precise sense that the per-coordinate $(\alpha,\varepsilon)$ R\'enyi divergence between the
output laws on two single-participation neighbours is bounded by that of one Gaussian mechanism of
multiplier $\sigma_{\mathrm{eff}}$, with equality for the worst-case neighbour:
$D_\alpha\big(M(D)\,\|\,M(D')\big)\le \alpha/(2\sigma_{\mathrm{eff}}^2)$ for all $\alpha>1$.
\item \textnormal{(Calibration.)} Consequently, to obtain the same R\'enyi/RDP guarantee (hence the
same $(\varepsilon,\delta)$ after conversion) as an i.i.d.\ DP\nobreakdash-SGD release with target noise
multiplier $\sigma$, it suffices to set
\begin{equation}
\boxed{\;\nu \;=\; \sigma\, C\, \kappa
\;=\;\sigma\,C\,\sqrt{\tfrac{1-\lambda^{2T}}{1-\lambda^{2}}}\; }.
\label{eq:calib}
\end{equation}
\item \textnormal{(Reduction at $\lambda=0$.)} For $\lambda=0$ one has $L=I$, $\kappa=1$, $n_t=z_t$,
and \eqref{eq:calib} gives $\nu=\sigma C$: the mechanism is exactly single-participation DP-SGD with
multiplier $\sigma$.
\item \textnormal{(The naive sensitivity is wrong.)} The factor is $\kappa=\|(L^{-1})_{:,t^\star}\|_2$,
\emph{not} the column norm $\sqrt{1+\lambda^2}=\|L_{:,t}\|_2$ of the injection matrix $L$. Using
$\sqrt{1+\lambda^2}$ under-scales the noise (e.g.\ $1.35$ vs.\ the correct $2.29$ at $\lambda=0.9$)
and \emph{violates} the intended DP guarantee.
\end{enumerate}
Finally, the correlation in \eqref{eq:inject} couples the per-step outputs, so Poisson-subsampling
privacy amplification no longer applies; the guarantee above is for the unamplified
single-participation accountant.
\end{theorem}

\begin{proof}
Throughout we exploit that both the injection \eqref{eq:inject} and the prefix-sum aggregation act
\emph{coordinate-wise and identically} on the $d$ coordinates of $\mathbb R^d$, with independent noise
across coordinates. Hence we may fix one coordinate, work with scalar streams in $\mathbb R^{T}$, and
recombine at the end via the Kronecker structure of the joint covariance (Step~2). The neighbour
relation is the standard add/remove-one-record relation; under single participation a neighbouring
dataset differs in the contents of \emph{exactly one} batch $\mathcal B_{t^\star}$, so the clipped-sum
stream $s=(s_1,\dots,s_T)$ changes only in the single block $s_{t^\star}$.

\medskip
\noindent\textbf{Step 0: Notation and the factorisation $A=B\,L^{-1}$ with $B=AL$.}
Stack the scalar (fixed-coordinate) injections as $z=(z_1,\dots,z_T)^\top$ (with $z_0=0$) and
$n=(n_1,\dots,n_T)^\top$. Equation \eqref{eq:inject} reads $n=Lz$, where $L\in\mathbb R^{T\times T}$ is
the lower-bidiagonal Toeplitz matrix
\[
L=\begin{pmatrix}
1 & & & \\
-\lambda & 1 & & \\
 & -\lambda & 1 & \\
 & & \ddots & \ddots \\
 & & & -\lambda & 1
\end{pmatrix},
\qquad
z\sim\mathcal N(0,\nu^2 I_T).
\]
Indeed $(Lz)_t=L_{t,t}z_t+L_{t,t-1}z_{t-1}=z_t-\lambda z_{t-1}=n_t$. $L$ is lower triangular with
unit diagonal, hence invertible with $\det L=1$. Its inverse is the lower-triangular Toeplitz matrix
of geometric weights,
\begin{equation}
(L^{-1})_{t,k}=\begin{cases}\lambda^{\,t-k}, & k\le t,\\ 0,&k>t,\end{cases}
\label{eq:Linv}
\end{equation}
which one verifies directly: for $k<t$,
$\big(L L^{-1}\big)_{t,k}=\sum_j L_{t,j}(L^{-1})_{j,k}
=(L^{-1})_{t,k}-\lambda (L^{-1})_{t-1,k}=\lambda^{t-k}-\lambda\cdot\lambda^{t-1-k}=0$;
for $k=t$ it equals $(L^{-1})_{t,t}=1$; for $k>t$ both terms vanish; and the boundary row $t=1$ gives
$(L^{-1})_{1,1}=1$. (Numerically, materialising $L$ and computing $L^{-1}$ reproduces \eqref{eq:Linv}
to machine precision, $\max$-error $\sim10^{-16}$.)

The DP-CorrMom mechanism outputs, per coordinate, the noisy summed-gradient stream
$\hat g=s+n=s+Lz$. The quantity that drives learning (and whose privacy we audit, since prefix sums
are post-processing of $\hat g$) is the running prefix sum $A\hat g$, $A$ lower-triangular all-ones.
Writing $z=\nu\zeta$ with $\zeta\sim\mathcal N(0,I_T)$,
\begin{equation}
A\hat g \;=\; A s \;+\; A L z \;=\; \underbrace{A s}_{\text{signal}}
\;+\;\nu\,\underbrace{(AL)}_{=:B}\,\zeta .
\label{eq:factor}
\end{equation}
Thus the released (post-processed) stream is the prefix-sum signal $As$ plus correlated Gaussian noise
with covariance $\nu^2 BB^\top$, $B=AL$. This is precisely the matrix-mechanism factorisation of the
workload $A$ through the \emph{strategy} $C_{\mathrm{strat}}:=L^{-1}$ and \emph{reconstruction}
$B:=AL$, since $B\,C_{\mathrm{strat}}=(AL)\,L^{-1}=A$. (Caution: the relevant matrices are the
reconstruction $AL$ acting on $\zeta$ in \eqref{eq:factor} and the strategy $L^{-1}$ whose column
norms set the sensitivity in Step~2; the product $A L^{-1}$ plays no role here.) This proves the
factorisation claim in (i); it remains to compute the privacy of the single Gaussian mechanism
$\zeta\mapsto \nu B\zeta$ added to $As$.

\medskip
\noindent\textbf{Step 1: The whole release is one Gaussian mechanism (Mahalanobis form).}
Condition on the data; the only randomness is $\zeta$. From \eqref{eq:factor}, per coordinate,
\begin{equation}
A\hat g \;\sim\; \mathcal N\!\big(As,\ \nu^2 B B^\top\big),\qquad B=AL .
\label{eq:law}
\end{equation}
$B=AL$ is a product of invertible matrices ($A$ is lower-triangular with unit diagonal, $\det A=1$),
hence invertible and the Gaussian \eqref{eq:law} is non-degenerate. Because $\hat g\mapsto A\hat g$ is
an invertible linear post-processing, releasing the prefix sum $A\hat g$ and releasing the raw stream
$\hat g=s+Lz\sim\mathcal N(s,\nu^2 LL^\top)$ are \emph{equivalent up to post-processing} and have
identical privacy; we phrase the computation on $\hat g$, whose law is
\begin{equation}
\hat g\sim\mathcal N\!\big(s,\ \Sigma\big),\qquad \Sigma:=\nu^2 LL^\top .
\label{eq:rawlaw}
\end{equation}
Equation \eqref{eq:rawlaw} is a \emph{single} multivariate Gaussian over the entire $T$-step output,
with a fixed (data-independent) covariance $\Sigma$ and a data-dependent mean $s$. Two neighbouring
datasets $D,D'$ produce the same $\Sigma$ and means $s,s'$ differing only in block $t^\star$. The
privacy of such a fixed-covariance, shifted-mean Gaussian family is governed by the Mahalanobis norm
of the mean shift, made precise next.

\begin{lemma}[R\'enyi divergence of a fixed-covariance Gaussian mechanism]
\label{lem:gauss-renyi}
Let $\Sigma\succ0$ and $P=\mathcal N(s,\Sigma)$, $Q=\mathcal N(s',\Sigma)$. Then for every
$\alpha>1$,
\begin{equation}
D_\alpha(P\,\|\,Q)\;=\;\frac{\alpha}{2}\,(s-s')^\top \Sigma^{-1}(s-s')
\;=\;\frac{\alpha}{2}\,\big\|s-s'\big\|_{\Sigma^{-1}}^{2}.
\label{eq:renyi}
\end{equation}
\end{lemma}
\begin{proof}
Whiten by $W:=\Sigma^{-1/2}$: $WP=\mathcal N(Ws,I)$ and $WQ=\mathcal N(Ws',I)$, and R\'enyi
divergence is invariant under the invertible map $W$. For isotropic unit-covariance Gaussians the
divergence factorises over coordinates and, along the shift direction $\Delta:=W(s-s')$, reduces to
the one-dimensional identity $D_\alpha\big(\mathcal N(0,1)\,\|\,\mathcal N(\delta,1)\big)
=\tfrac{\alpha}{2}\delta^2$ (a Gaussian integral that converges for all $\alpha>1$, with the
covariance-mismatch terms of the general Gaussian R\'enyi formula absent because the two covariances
coincide). Hence $D_\alpha(P\|Q)=\tfrac{\alpha}{2}\|\Delta\|_2^2=\tfrac{\alpha}{2}(s-s')^\top\Sigma^{-1}(s-s')$.
\end{proof}

\noindent
(We verified \eqref{eq:renyi} numerically: for scalar shifts the closed form matches direct numerical
integration of $\tfrac1{\alpha-1}\log\!\int p^\alpha q^{1-\alpha}$ for $\alpha\in\{2,3,8\}$, and for the
non-diagonal $\Sigma=\nu^2LL^\top$ the multivariate value $\tfrac\alpha2(s-s')^\top\Sigma^{-1}(s-s')$
equals $\tfrac{\alpha}{2\nu^2}\|L^{-1}(s-s')\|_2^2$ exactly.)

Apply Lemma~\ref{lem:gauss-renyi} with $\Sigma=\nu^2 LL^\top$ and $\Delta s:=s-s'$ supported on block
$t^\star$. Since $\Sigma^{-1}=\nu^{-2}(LL^\top)^{-1}=\nu^{-2}L^{-\top}L^{-1}$,
\begin{equation}
D_\alpha\big(P\|Q\big)
=\frac{\alpha}{2\nu^{2}}\,\Delta s^\top L^{-\top}L^{-1}\,\Delta s
=\frac{\alpha}{2\nu^{2}}\,\big\|L^{-1}\Delta s\big\|_2^{2}.
\label{eq:renyi-mahal}
\end{equation}
This is exactly the R\'enyi divergence of a \emph{single} Gaussian mechanism with sensitivity
$\|L^{-1}\Delta s\|_2$ and scale $\nu$: it certifies the entire $T$-step correlated release at once,
not step-by-step.

\medskip
\noindent\textbf{Step 2: Worst-case sensitivity is $C\kappa$ with $\kappa=\operatorname{maxcol}(L^{-1})$.}
We first establish the $d$-coordinate Mahalanobis sensitivity. The joint covariance over the
$Td$-dimensional output is $\Sigma_{\mathrm{full}}=\nu^2\,(LL^\top)\otimes I_d$ (the same temporal $L$
per coordinate, with noise independent across coordinates). Under single participation a one-record
change perturbs only block $t^\star$, so the mean shift is $\Delta=e_{t^\star}\otimes u$ with
$u\in\mathbb R^d$ the per-coordinate change of the clipped sum and $e_{t^\star}\in\mathbb R^T$ a
standard basis vector. Using $(P\otimes Q)^{-1}=P^{-1}\otimes Q^{-1}$ and
$(a\otimes u)^\top(P\otimes I)(a\otimes u)=(a^\top P a)(u^\top u)$,
\begin{equation}
\Delta^\top\Sigma_{\mathrm{full}}^{-1}\Delta
=\nu^{-2}\big(e_{t^\star}^\top(LL^\top)^{-1}e_{t^\star}\big)\,\|u\|_2^2
=\nu^{-2}\,\big\|(L^{-1})_{:,t^\star}\big\|_2^2\,\|u\|_2^2,
\label{eq:kron}
\end{equation}
where the last identity uses $(LL^\top)^{-1}_{tt}=\big(L^{-\top}L^{-1}\big)_{tt}=\|(L^{-1})_{:,t}\|_2^2$.
(We verified \eqref{eq:kron} numerically against the dense $Td\times Td$ inverse.) By the
add/remove-one-record relation and per-sample clipping at norm $C$, the full change in the clipped sum
at step $t^\star$ is a single clipped per-sample gradient, $\|u\|_2=\|\operatorname{clip}_C(\nabla\ell_{i^\star})\|_2\le C$,
with all other blocks unchanged. Maximising over which batch is perturbed gives the worst-case
$\ell_2$ Mahalanobis sensitivity of the whole release,
\begin{equation}
\Delta_2 \;:=\; \max_{D\sim D'}\big\|L^{-1}\Delta s\big\|_2
\;=\; C\cdot \max_{1\le t\le T}\big\|(L^{-1})_{:,t}\big\|_2
\;=\; C\,\kappa ,
\label{eq:sens}
\end{equation}
attained when $\|u\|_2=C$. It remains to evaluate $\kappa$. By \eqref{eq:Linv} the $t$-th column of
$L^{-1}$ has entries $(L^{-1})_{r,t}=\lambda^{\,r-t}$ for $r\ge t$ and $0$ otherwise, so its squared
norm is a finite geometric sum over the $T-t+1$ rows on or below the diagonal:
\begin{equation}
\big\|(L^{-1})_{:,t}\big\|_2^2=\sum_{r=t}^{T}\lambda^{2(r-t)}=\sum_{k=0}^{T-t}\lambda^{2k}
=\frac{1-\lambda^{2(T-t+1)}}{1-\lambda^{2}}\qquad(\lambda\in(0,1)).
\label{eq:colnorm}
\end{equation}
This is strictly decreasing in $t$ (each later column has fewer nonzero entries), so the maximum is at
$t=1$ (the earliest step has the longest tail of geometric weights), giving
\begin{equation}
\kappa=\big\|(L^{-1})_{:,1}\big\|_2=\sqrt{\sum_{k=0}^{T-1}\lambda^{2k}}
=\sqrt{\frac{1-\lambda^{2T}}{1-\lambda^{2}}},
\label{eq:kappa-cf}
\end{equation}
which is exactly \texttt{corr\_sensitivity}$(\lambda,T)=\sqrt{(1-\lambda^{2T})/(1-\lambda^2)}$ in
\texttt{dp\_adaptive.py}. For $\lambda\to0$, $\kappa=1$; as $T\to\infty$, $\kappa\to1/\sqrt{1-\lambda^2}$.
(Numerically, $\max_t\|(L^{-1})_{:,t}\|_2$ and $\sqrt{\max_t(LL^\top)^{-1}_{tt}}$ both equal
\eqref{eq:kappa-cf} to machine precision, with the argmax at $t=1$; e.g.\ $\kappa=2.2942$ at
$\lambda=0.9$ for large $T$.)

\medskip
\noindent\textbf{Step 3: Single effective multiplier.}
Substituting \eqref{eq:sens} into \eqref{eq:renyi-mahal}, the entire $T$-step DP-CorrMom release
satisfies, for all $\alpha>1$,
\begin{equation}
D_\alpha\big(M(D)\,\|\,M(D')\big)\;\le\;\frac{\alpha}{2}\cdot\frac{\Delta_2^2}{\nu^2}
=\frac{\alpha}{2}\cdot\frac{C^2\kappa^2}{\nu^2}
=\frac{\alpha}{2\,\sigma_{\mathrm{eff}}^2},
\qquad
\sigma_{\mathrm{eff}}:=\frac{\nu}{C\kappa},
\label{eq:final-renyi}
\end{equation}
with equality for the worst-case neighbour (perturb batch $t^\star=1$ with a record whose clipped
gradient has norm exactly $C$ aligned with the shift; by \eqref{eq:sens} this attains $\Delta_2=C\kappa$
and Lemma~\ref{lem:gauss-renyi} is an equality). The right-hand side is precisely the R\'enyi
divergence of one Gaussian mechanism of sensitivity $1$ and noise multiplier $\sigma_{\mathrm{eff}}$.
Hence the whole correlated, $T$-step release is, for RDP accounting, a \emph{single} Gaussian mechanism
with effective multiplier $\sigma_{\mathrm{eff}}=\nu/(C\kappa)$, proving claim (i).

\medskip
\noindent\textbf{Step 4: Calibration $\nu=\sigma C\kappa$ (claim (ii)).}
An i.i.d.\ DP-SGD release with target multiplier $\sigma$ is the Gaussian mechanism with
$\sigma_{\mathrm{eff}}=\sigma$ (the case $\lambda=0$, $\kappa=1$, $\nu=\sigma C$, where each step adds
$\mathcal N(0,(\sigma C)^2 I)$ to a sum of $C$-clipped gradients). By \eqref{eq:final-renyi} the
DP-CorrMom release has the \emph{same} R\'enyi bound $\alpha/(2\sigma_{\mathrm{eff}}^2)$ for every order
$\alpha$ iff $\sigma_{\mathrm{eff}}=\sigma$, i.e.
\[
\frac{\nu}{C\kappa}=\sigma
\quad\Longleftrightarrow\quad
\nu=\sigma\,C\,\kappa=\sigma\,C\,\sqrt{\frac{1-\lambda^{2T}}{1-\lambda^2}} .
\]
Because the RDP curve $\alpha\mapsto \alpha/(2\sigma^2)$ is identical to that of i.i.d.\ DP-SGD for
\emph{all} $\alpha$, any RDP/zCDP$\to(\varepsilon,\delta)$ conversion (e.g.\ the PRV/Opacus accountant
used in the implementation) yields the same $(\varepsilon,\delta)$. This is exactly the rule the caller
implements: inflate the per-step base std by $\kappa$
($\texttt{noise\_multiplier}\mathrel{*}{=}\texttt{corr\_sensitivity}(\lambda,T)$) and run the
\emph{unamplified} accountant at $\sigma$. This proves (ii).

\medskip
\noindent\textbf{Step 5: $\lambda=0$ reduces to single-participation DP-SGD (claim (iii)).}
At $\lambda=0$, $L=I_T$, so $\Sigma=\nu^2 I_T$, $L^{-1}=I_T$, $\kappa=\|e_t\|_2=1$, and the injection
\eqref{eq:inject} is $n_t=z_t\sim\mathcal N(0,\nu^2 I_d)$, i.i.d.\ across $t$. The calibration
\eqref{eq:calib} collapses to $\nu=\sigma C$ and \eqref{eq:final-renyi} to $D_\alpha\le\alpha/(2\sigma^2)$,
the single-participation DP-SGD bound: the composition of $T$ \emph{disjoint-batch} Gaussian mechanisms
equals one Gaussian mechanism because, with disjoint batches, each record touches exactly one step
(parallel, not sequential, composition). The implementation confirms this bit-for-bit: with
$\lambda=0$, $\texttt{\_inject\_correlated\_noise}$ sets $w=z$, reproducing stock i.i.d.\ Opacus noise.
This proves (iii).

\medskip
\noindent\textbf{Step 6: The naive $\sqrt{1+\lambda^2}$ is wrong (claim (iv)).}
A tempting but incorrect shortcut reads the sensitivity off the \emph{injection} matrix $L$ rather than
off the \emph{strategy} $L^{-1}$. The interior columns of $L$ have norm
\[
\big\|L_{:,t}\big\|_2=\sqrt{L_{t,t}^2+L_{t+1,t}^2}=\sqrt{1+\lambda^2}\qquad(1\le t\le T-1),
\]
so the naive recipe would set $\nu_{\mathrm{naive}}=\sigma C\sqrt{1+\lambda^2}$. This is the wrong
object: in \eqref{eq:renyi-mahal} the matrix that appears is
$\Sigma^{-1}=\nu^{-2}L^{-\top}L^{-1}$, whose diagonal entries are
$(LL^\top)^{-1}_{tt}=\|(L^{-1})_{:,t}\|_2^2$, \emph{not} $\|L_{:,t}\|_2^2$. Equivalently, a one-record
change perturbs the \emph{released} stream through $L^{-1}$, bringing in the geometric tail
$1,\lambda,\lambda^2,\dots$, not the two-term stencil $1,-\lambda$. Numerically, at $\lambda=0.9$,
\[
\sqrt{1+\lambda^2}=1.345\quad\text{vs.}\quad
\kappa=\sqrt{\tfrac{1-\lambda^{2T}}{1-\lambda^2}}\xrightarrow[T\to\infty]{}\frac{1}{\sqrt{1-\lambda^2}}=2.294 .
\]
Using $\nu_{\mathrm{naive}}=1.345\,\sigma C<2.294\,\sigma C=\nu$ under-scales the noise by a factor
$\kappa/\sqrt{1+\lambda^2}\approx1.71$, giving an effective multiplier
$\sigma_{\mathrm{eff}}^{\mathrm{naive}}=\nu_{\mathrm{naive}}/(C\kappa)=\sigma\sqrt{1+\lambda^2}/\kappa
\approx 0.586\,\sigma<\sigma$, a strictly weaker (\emph{violated}) guarantee: by \eqref{eq:final-renyi}
the true R\'enyi divergence is $\alpha/(2(\sigma_{\mathrm{eff}}^{\mathrm{naive}})^2)
=\alpha\kappa^2/(2(1+\lambda^2)\sigma^2)$, inflated by $\kappa^2/(1+\lambda^2)\approx2.9$ over the
claimed bound. Hence the correct sensitivity is $\kappa=\operatorname{maxcol}(L^{-1})$ and the naive
$\sqrt{1+\lambda^2}$ breaks privacy. This proves (iv).

\medskip
\noindent\textbf{Step 7: Amplification is voided.}
Privacy amplification by Poisson subsampling requires the per-step mechanisms to be \emph{independent}
given the data, so that the subsampling-induced mixture tightens each step's divergence and composition
multiplies the tightened terms. Here the injected noises $n_t=z_t-\lambda z_{t-1}$ share the variable
$z_{t-1}$ across consecutive steps; the joint law \eqref{eq:rawlaw} has the non-diagonal covariance
$\Sigma=\nu^2 LL^\top$ (off-diagonal $-\lambda\nu^2$ on the first sub/super-diagonal), so the steps are
statistically coupled and the amplification argument's independence hypothesis fails. We therefore
account the release as a single (unamplified) Gaussian mechanism via \eqref{eq:final-renyi}, with single
participation ensuring each record's contribution is the single column $(L^{-1})_{:,t^\star}$ used in
Step~2. (Multi-epoch participation would activate several columns of $L^{-1}$ for one record and break
the single-column sensitivity bound; it requires a min-separation or banded analysis and is outside
this statement.) This establishes the final clause and completes the proof.
\end{proof}

\par\smallskip\noindent\textbf{Remark (consistency with the utility intuition).}\quad 
The same $\kappa$ that inflates the injected noise is what makes the \emph{integrated} (prefix-sum)
noise small. The released noise on the prefix sum has covariance $\nu^2 BB^\top=\nu^2 ALL^\top A^\top$;
at matched $(\varepsilon,\delta)$ (so $\nu=\sigma C\kappa$) its per-output variance and the
i.i.d.\ DP-SGD per-output variance \emph{both} grow as $\Theta(T)$, but the correlated mechanism wins by
a \emph{constant factor}: the asymptotic ratio of the prefix-sum noise variance versus i.i.d.\ DP-SGD
tends to $(1-\lambda)/(1+\lambda)<1$ as $T\to\infty$ (e.g.\ $\approx0.053$ at $\lambda=0.9$; verified
numerically). This is the provable variance reduction that motivates DP-CorrMom. The empirical finding
of this project is that at the operating point $\rho\approx1$ (second moment a noise floor, signal below
the floor) this variance reduction does not translate into a utility gain, because the model is
signal-limited rather than variance-limited.